\chapter*{Appendices}
\addcontentsline{toc}{chapter}{Appendices}

\section*{Citation Registry}
The master list of approved citations for the Catty framework.

\section*{DSL Reference}
Full specification for the Catty Domain-Specific Language.

\section*{SPARQL Query Reference}

\subsection*{External SPARQL Queries (Executable)}

This section documents SPARQL queries that have been manually tested against external SPARQL endpoints using curl. All results are actual endpoint outputs with zero LLM generation.

\subsubsection*{Q1: Wikidata QID Verification}

\textbf{Purpose}: Verify that a QID matches its intended concept.

\textbf{Endpoint}: \code{https://query.wikidata.org/sparql}

\textbf{Query}:
\begin{lstlisting}[language=SPARQL,breaklines=true]
PREFIX wd: <http://www.wikidata.org/entity/>
PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
PREFIX schema: <http://schema.org/>

SELECT ?label ?description WHERE {
  wd:Q193138 rdfs:label ?label .
  wd:Q193138 schema:description ?description .
  FILTER(LANG(?label) = "en")
}
\end{lstlisting}

\textbf{Result}: Q193138 = ``Les Mars'' (French commune)

\textbf{Analysis}: This demonstrates the QID hallucination problem. Q193138 is NOT ``natural transformation'', validating the warning against guessing QIDs.

\textbf{Status}: VALID -- Execution time: 0.754s, Result count: 1 item

\subsubsection*{Q2: Wikidata Label Discovery}

\textbf{Purpose}: Find the correct QID for a concept by label.

\textbf{Endpoint}: \code{https://query.wikidata.org/sparql}

\textbf{Query}:
\begin{lstlisting}[language=SPARQL,breaklines=true]
PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
PREFIX schema: <http://schema.org/>

SELECT DISTINCT ?item ?label ?description WHERE {
  ?item rdfs:label "natural transformation"@en .
  BIND("natural transformation"@en AS ?label)
  ?item schema:description ?description .
  FILTER(LANG(?description) = "en")
}
LIMIT 5
\end{lstlisting}

\textbf{Result}: Q1442189 = ``natural transformation'' (transformation between two functors studied in category theory)

\textbf{Analysis}: Successfully discovers the correct QID for ``natural transformation'', matching the verified QID in the Core Domain Registry.

\textbf{Status}: VALID -- Execution time: 0.367s, Result count: 1 binding

\subsubsection*{Q3: Wikidata Logics Query}

\textbf{Purpose}: Find all logics in Wikidata that are instances of ``logic'' (Q8078).

\textbf{Endpoint}: \code{https://query.wikidata.org/sparql}

\textbf{Query}:
\begin{lstlisting}[language=SPARQL,breaklines=true]
PREFIX wd: <http://www.wikidata.org/entity/>
PREFIX wdt: <http://www.wikidata.org/prop/direct/>
PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>

SELECT ?logic ?logicLabel ?sequentForm
WHERE {
  ?logic wdt:P31 wd:Q8078 .  # instance of logic
  ?logic rdfs:label ?logicLabel .
  FILTER(LANG(?logicLabel) = "en")
  OPTIONAL { ?logic wdt:P361 ?sequentForm . }
}
LIMIT 20
\end{lstlisting}

\textbf{Results}: 5 logic instances found:
\begin{itemize}
    \item Q7782444 -- ``Theory of obligationes''
    \item Q28171908 -- ``Jaina seven-valued logic''
    \item Q65122115 -- ``game semantics''
    \item Q107417121 -- ``strict logic''
    \item Q113995870 -- ``elimination method''
\end{itemize}

\textbf{Analysis}: Successfully retrieves logics from Wikidata. The OPTIONAL clause for sequentForm returned no results, indicating Wikidata lacks sequent form information.

\textbf{Status}: VALID -- Execution time: 0.939s, Result count: 5 bindings

\subsection*{Queries Requiring Refactoring to External Standards}

The following 13 queries in \code{docs/sparql-examples.md} use \textbf{custom vocabulary that must be replaced} with W3C/OMG/ISO standards to comply with AGENTS.md constraints on consuming external semantic web data rather than authoring new ontologies.

\subsubsection*{Query Refactoring Requirements}

\begin{enumerate}
    \item \textbf{Basic Queries}: \code{all-logics.rq}, \code{structural-rules.rq}
    \begin{itemize}
        \item Replace \code{catty:Logic} with \code{owl:Ontology} (W3C OWL 2)
        \item Replace \code{catty:hasLogicalAxiom} with \code{owl:Axiom} (W3C OWL 2)
    \end{itemize}
    
    \item \textbf{Morphism Queries}: \code{morphisms-from-logic.rq}, \code{lattice-order.rq}
    \begin{itemize}
        \item Replace \code{catty:domain/codomain} with \code{theory:domain/codomain} (ISO Common Logic/COLORE)
        \item Replace \code{catty:Extension} with COLORE Category Theory morphism structures
    \end{itemize}
    
    \item \textbf{Adjoint Queries}: \code{adjoint-relationships.rq}
    \begin{itemize}
        \item Replace \code{catty:AdjointFunctors} with COLORE Category Theory adjunction formalization
    \end{itemize}
    
    \item \textbf{Curry-Howard Queries}: \code{curry-howard-mapping.rq}, \code{curry-howard-functor.rq}
    \begin{itemize}
        \item Replace \code{catty:CurryHowardFunctor} with \code{dol:LogicTranslation} (OMG DOL)
        \item Replace \code{catty:correspondsToLogic} with DOL logic mapping predicates
    \end{itemize}
    
    \item \textbf{Lattice Queries}: \code{extension-hierarchy.rq}, \code{lattice-neighbors.rq}
    \begin{itemize}
        \item Refactor to use COLORE partial order structures from Common Logic
    \end{itemize}
    
    \item \textbf{Validation Queries}: \code{orphan-logics.rq}, \code{terminal-logics.rq}
    \begin{itemize}
        \item Refactor to validate against OWL 2 ontology completeness constraints
    \end{itemize}
    
    \item \textbf{Complex Queries}: \code{find-all-paths-in-lattice.rq}, \code{validate-curry-howard-correspondence.rq}
    \begin{itemize}
        \item Refactor to use DOL logic translation validation and COLORE morphism composition
    \end{itemize}
\end{enumerate}

\textbf{Action Required}: All queries must be refactored to import and consume from W3C OWL 2, OMG DOL, and ISO COLORE repositories rather than defining new \code{catty:} vocabulary.

\section*{SHACL Shape Reference}
Validation constraints for the formal ontology.
