% Theorem: Exchange as Permutation Invariance
% ID: thm-exchange-permutation
% Research Source: DBpedia - Structural proof theory

\input{../../../../../thm-preamble}

\title{Theorem: Exchange as Permutation Invariance}
\date{}

\begin{document}

\thmid{exchange-permutation}

\begin{theorem}[Structural Exchange as Permutation Invariance]
\label{thm:exchange-permutation}
In classical sequent calculus (LK), the exchange rule provides complete permutation invariance: if $\Gamma \vdash \Delta$ is derivable, then $\Gamma' \vdash \Delta'$ is derivable for any permutations $\Gamma'$ of $\Gamma$ and $\Delta'$ of $\Delta$.
\end{theorem}

\begin{proof}
We prove by induction on the length of the derivation.

\textbf{Base Case:} Axiom rule $\overline{A \vdash A}$
Any permutation of $A \vdash A$ is identical to itself, so the property holds trivially.

\textbf{Inductive Step:} Consider a rule application in the derivation.

For left rules, suppose the last rule is:
\[
\infer[\land L_i]{\Gamma, A \land B, \Delta \vdash \Sigma}{\Gamma, A, \Delta \vdash \Sigma}
\]

By the induction hypothesis, $\Gamma, A, \Delta \vdash \Sigma$ is invariant under permutation of its antecedent. The exchange rules allow us to permute $A \land B$ with any formula in $\Gamma$ or $\Delta$, establishing that the conclusion is also permutation invariant.

For right rules, the argument is symmetric using right exchange.

The key insight is that the exchange rules:
\[
\infer[\text{Exchange}_L]{\Gamma, B, A, \Delta \vdash \Sigma}{\Gamma, A, B, \Delta \vdash \Sigma}
\quad
\infer[\text{Exchange}_R]{\Gamma \vdash \Delta, B, A, \Sigma}{\Gamma \vdash \Delta, A, B, \Sigma}
\]
generate the full symmetric group on formula positions. Since any permutation can be decomposed into adjacent transpositions, and exchange provides exactly these transpositions, complete permutation invariance follows.
\end{proof}

\begin{corollary}[Multiset Semantics]
The exchange rule makes sequent contexts behave as multisets (where order does not matter) rather than sequences or lists.
\end{corollary}

\begin{remark}[Non-Commutative Variants]
When exchange is rejected:
\begin{itemize}
\item \textbf{Lambek Calculus:} Sequents are ordered; formula position matters semantically.
\item \textbf{Bunched Implications:} Combines commutative and non-commutative contexts using tree structures.
\item \textbf{Categorical Model:} Without exchange, categories are monoidal but not necessarily symmetric.
\end{itemize}
\end{remark}

\begin{remark}[Duality]
The duality between left and right exchange mirrors the symmetry of the sequent arrow. In category theory, this corresponds to the symmetry isomorphism $\sigma_{A,B} : A \otimes B \cong B \otimes A$ in symmetric monoidal categories.
\end{remark}

\end{document}
